Automatically generated NE tagged corpora for English and Hungarian

### 2012-03-25 paper/engNERwiki.tex
ez

A NER corpus consists of sentences where NEs are accurately tagged. In order to automatically prepare such sentences, two tasks need to be performed: identifying entities in the sentence and tagging them with the correct tag. Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.

@@

There are some special cases to the algorithm described thus far. These are detailed below.

@@

For some of the well-resourced languages, DBpedia has internationalized "chapters" -- data dumps based on the Wikipedia of those particular languages. Hungarian does not have a separate chapter. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.

@@

The files are in multitag format. Content lines are tab separated; there is one column for the tokens plus one column per tagset. Sentence boundaries are marked by empty lines. The linguistic features include the lemmatized form of the word and its part-of-speech tag. Two NE tags are included with each word: one is the most specific DBpedia category it belongs to, if any; the other is the CoNLL NE tag. While the NE tags can be considered as a "silver standard", the linguistic features are provided on a "best-effort" basis.

@@

Our automatic annotation process retains all of the Wikipedia sentences, which remained after our two-fold filtering method. Thus sentences without NEs are also included in the corpus. The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible. However, after further investigation of the NE density in our Wikipedia corpora and gold standard corpora (see Table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.

@@

For English cross-corpus evaluation the CoNLL-2003 dataset was chosen, mostly because of the similarity in tagset and format. It is well-studied that cross-checking between different corpora decrease F-measure. Stylistic and domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems. Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs much better on corresponding test sets than on test sets from other sources. The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own training set (see Table \ref{enresults} for results).

However, Wikipedia-derived corpora can also be used for improving NER performance in other ways. First, like e.g.~\cite{Toral:06} we collected huge gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset. A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train. Both methods result in improved overall F-measure on CoNLL test set.

Since in NE labeling for Hungarian we followed the Szeged NER corpus annotation guidelines, it is obvious to perform the experiments on this dataset, as well. Hungarian results are quite similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher, than ones for English. It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency and diversity of this class is smaller (see section \ref{sec:hun}). In contrast to the CoNLL corpus, the Szeged NER corpus was highly accurately annotated with an inter-annotator agreement over 99\%.

Due to the quite good F-measure of training on the Hungarian Wikipedia train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE-taggers for domains falling out of newswire.

@@

Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, like Hungarian, as well. Due to the particularities of the Hungarian language, some steps are slightly different, and special linguistics phenomena pose several problems related to the NER task to solve.

Automatically generated, silver standard corpora can be very useful for improving NER in more ways. We showed that huge gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves overall F-measure. Moreover, our Hungarian corpus can serve as a training corpus to build NE taggers for more general domains, than the classic newswire.

>>>>

In order to automatically prepare sentences where NEs are accurately tagged, two tasks need to be performed: identifying entities in the sentence and tagging them with the correct tag. Sentences, for which accurate tagging could not be accomplished, must be removed from the corpus.

@@

%There are some special cases to the method, which are detailed below.

@@

For well-resourced languages, DBpedia has internationalized chapters; data dumps based on Wikipedia. Hungarian does not have a separate chapter. Instead, the Hungarian entity list comprises of the pages in the English list that have their equivalents in the Hungarian Wikipedia. Two consequences follow.

@@

The files are in multitag format. Content lines are tab separated; there is one column for the tokens plus one column per tagset. Sentence boundaries are marked by empty lines. The linguistic features include the lemmatized form of the word and its part-of-speech tag. Two NE tags are included with each word: the most specific DBpedia category it belongs to and the CoNLL NE tag. While the NE tags can be considered as a "silver standard", the linguistic features are provided on a "best-effort" basis.

@@

Our automatic annotation process retains all of the Wikipedia sentences, which remained after our two-fold filtering method. Thus sentences without NEs are also included in the corpus. The rationale behind this is that we wanted to reserve the original distribution of names in Wikipedia as much as possible. However, after further investigation of the NE density in our corpora and gold standard corpora (see Table \ref{size}), we decided not to include the sentences without NEs in evaluation datasets.

@@

For English cross-corpus evaluation the CoNLL-2003 corpus was chosen. It is well-studied that cross-checking between different corpora decrease F-measure. Domain differences certainly affect NER performance, and the different annotation schemes pose several compatibility problems. Nothman et al.~\shortcite{Nothman:08} showed that each set of gold standard training data performs better on corresponding test sets than on test sets from other sources. The situation here is similar: the NE-tagger trained on Wikipedia does not achieve as high performance tested against CoNLL test set as one trained on the own train set (see Table \ref{enresults}).

However, Wikipedia-derived corpora can also be used for improving NER accuracy in other ways. First, we collected gazetteer lists from the corpus for each NE category, which improved the overall F-measure given to the NE-tagger training and testing on CoNLL dataset. A second trial was labeling the CoNLL datasets by the model trained on Wikipedia corpus, and giving these labels as extra features to the next CoNLL train. Both methods result in improved F-measure on CoNLL test set.

Since in Hungarian NE tagging we followed the Szeged NER corpus annotation guidelines, we performed the experiments on this dataset. Hungarian results are similar to the English ones (see Table \ref{huresults}), the only difference is that F-measures for Hungarian are significantly higher. It can be due to the fact that the MISC category for Hungarian contains less types of names, thus the inconsistency of this class is smaller (see section \ref{sec:hun}). In contrast to the CoNLL corpus, the Szeged NER corpus was accurately annotated with an inter-annotator agreement over 99\%.

Due to the quite good F-measure of training on our Hungarian train corpus and testing on the corresponding test set, our Hungarian corpus can serve as a training corpus to build NE-taggers for domains falling out of newswire.

@@

Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, like Hungarian, as well. Due to the particularities of Hungarian, some steps are slightly different, and special linguistics phenomena pose several problems related to the NER task to solve.

Automatically generated corpora can be useful for improving NER in more ways. We showed that gazetteer lists extracted from our corpora, and training with extra features given by the model trained on our corpora, improves F-measure. Moreover, our Hungarian corpus can serve as a training corpus for more general domains. 
%than the classic newswire.
